### 2022-07-08 src/text/3-applications-subsections/3D-analysis.tex
Adapt application description

\subsection{3D Analysis}
\label{ssec:3d-analysis}
\todo{Laura}
The 1D analysis approach is a powerful tool to measure the spectrum of an
isolated source. However, more complicated situations require a more careful
treatment. In a field of view containing several overlapping sources, the 1D
approach can't disentangle the contribution of each source to the total flux in
the selected region. Sources with extended or complex morphology can result in
the measured flux being underestimated, and heavily dependent on the choice of
extraction region. Additionally, the 1D approach neglects the energy-dependence
of the PSF.

For such situations a more complex approach is needed, the so-called 3D
analysis. The three relevant dimensions are the two spatial angular coordinates
and an energy axis. In this framework, a combined spatial and spectral model
(that is, a SkyModel, see Section~\ref{ssec:gammapy-modeling}) is fitted to the
sky-maps that were previously derived from the data and bundled into a
MapDataset (see Sections~\ref{ssec:gammapy-makers} and~\ref{ssec:gammapy-datasets}).

@@

Starting from the \irfs corresponding to the same three simulated \cta
observations used in Section~\ref{ssec:1d-analysis}, we can create a MapDataset
via the MapDatasetMaker. However, we will not use the simulated event lists
provided by \cta but instead, use the method MapDataset.fake() to simulate
measured counts from the combination of several SkyModel instances. In this
way, a DL4 dataset can directly be simulated. In particular we simulate:
\begin{enumerate} \item A point source located at (l=0\textdegree, b=0\textdegree) with a power-law
	      spectral shape. \item An extended source with Gaussian morphology located at (l=0.4\textdegree,
	      b=0.15\textdegree) with $\sigma$=0.2\textdegree and a log-parabola spectral
	      shape. \item A large shell-like structure centered around (l=0.06\textdegree,
	      b=0.6\textdegree) with a radius and width of 0.6\textdegree and 0.3\textdegree
	      respectively and a power-law spectral shape. \end{enumerate} The position and sizes of the sources
have been selected so that their contributions overlap. This can be clearly
seen in the significance map shown in the left panel of
Figure~\ref{fig:cube_analysis}. This map was produced with the
ExcessMapEstimator (see Section~\ref{ssec:gammapy-estimators}) with a
correlation radius of 0.1\textdegree.

We can now fit the same model shapes to the simulated data and retrieve the
best-fit parameters. To check the model agreement, we compute the residual
significance map after removing the contribution from each model. This is done
again via the ExcessMapEstimator. As can be seen in the middle panel of
Figure~\ref{fig:cube_analysis}, there are no regions above or below 5$\sigma$,
meaning that the models describe the data sufficiently well.

>>>>

\subsection{3D Analysis}
\label{ssec:3d-analysis}
\todo{Laura}
The 1D analysis approach is a powerful tool to measure the spectrum of an
isolated source. However, more complicated situations require a more careful
treatment. In a field of view containing several overlapping sources, the 1D
approach cannot disentangle the contribution of each source to the total flux in
the selected region. Sources with extended or complex morphology can result in
the measured flux being underestimated, and heavily dependent on the choice of
extraction region. Additionally, the 1D approach neglects the energy-dependence
of the PSF.

For such situations a more complex approach is needed, the so-called 3D
analysis. The three relevant dimensions are the two spatial angular coordinates
and an energy axis. In this framework, a combined spatial and spectral model
(that is, a \code{SkyModel}, see Section~\ref{ssec:gammapy-modeling}) is fitted to the
sky-maps that were previously derived from the data and bundled into a
\code{MapDataset} (see Sections~\ref{ssec:gammapy-makers} and~\ref{ssec:gammapy-datasets}).

@@

Starting from the \irfs corresponding to the same three simulated \cta
observations used in Section~\ref{ssec:1d-analysis}, we can create a \code{MapDataset}
via the \code{MapDatasetMaker}. However, we will not use the simulated event lists
provided by \cta but instead, use the method MapDataset.fake() to simulate
measured counts from the combination of several SkyModel instances. In this
way, a DL4 dataset can directly be simulated. In particular we simulate:

\begin{enumerate}
    \item A point source located at (l=0\textdegree, b=0\textdegree) with a power-law
	      spectral shape.
    \item An extended source with Gaussian morphology located at (l=0.4\textdegree,
	      b=0.15\textdegree) with $\sigma$=0.2\textdegree and a log-parabola spectral
	      shape.
    \item A large shell-like structure centered around (l=0.06\textdegree,
	      b=0.6\textdegree) with a radius and width of 0.6\textdegree and 0.3\textdegree
	      respectively and a power-law spectral shape.
\end{enumerate}

The position and sizes of the sources
have been selected so that their contributions overlap. This can be clearly
seen in the significance map shown in the left panel of
Figure~\ref{fig:cube_analysis}. This map was produced with the
\code{ExcessMapEstimator} (see Section~\ref{ssec:gammapy-estimators}) with a
correlation radius of 0.1\textdegree.

We can now fit the same model shapes to the simulated data and retrieve the
best-fit parameters. To check the model agreement, we compute the residual
significance map after removing the contribution from each model. This is done
again via the \code{ExcessMapEstimator}. As can be seen in the middle panel of
Figure~\ref{fig:cube_analysis}, there are no regions above or below 5$\sigma$,
meaning that the models describe the data sufficiently well.

### 2022-07-08 src/text/3-applications-subsections/multi-instrument-analysis.tex
Adapt application description

The result of the joint fit is displayed in
Fig.~\ref{fig:multi_instrument_analysis}. We remark that the objective of this
exercise is illustrative, we display the flexibility of \gammapy in
simultaneously fitting multi-isntrument data even at different levels of
reduction, we do not aim to provide a new measurement of the Crab Nebula
spectrum.

>>>>

The result of the joint fit is displayed in
Fig.~\ref{fig:multi_instrument_analysis}. We remark that the objective of this
exercise is illustrative, we display the flexibility of \gammapy in
simultaneously fitting multi-instrument data even at different levels of
reduction, we do not aim to provide a new measurement of the Crab Nebula
spectrum.

### 2022-07-08 src/text/3-applications-subsections/temporal-analysis.tex
Adapt application description

A common use case in most astrophysical scenarios is to study the temporal
variability of the source. The most basic way to do this is to construct a
lightcurve, ie, the flux of the source in each given time bin. In gammapy, this
is done by using the LightCurveEstimator which fits the normalisation of the
source in each energy band per observation, keeping other parameters constant.
For custom time binning, an observation needs to be split into finer ones using
the \verb"Observation.select_time method". Figure \ref{fig:hess_lightcurve_pks}
shows the lightcurve of the blazar PKS~2155-304 in different energy bands as
observed by the H.E.S.S. telescope during an exceptional flare on the night of
July 29 - 30, 2006~\cite{2009A&A...502..749A}. The data is available publicly
as a part of the HESS-DL3-DR1~\cite{HESS-DL3-DR1}, and shipped with
\verb"GAMMAPY_DATA". Each observation is first split into 10 min smaller
observations, and spectra extracted for each of these within a 0.11 deg radius
around the source. A PowerLawSpectralModel is fit to all the datasets, leading
to a reconstructed index of $3.54 \pm 0.02$. With this assumed spectral model
the \verb"LightCurveEstimator" run directly for two energy bands, 0.5 - 1.5
TeV, and 1.5 - 20 TeV,	respectively.

@@

The obtained flux points can be analytically modelled using the available, or
user-implemented temporal models. Alternatively, instead of  extracting a
lightcurve, it is also possible to directly fit temporal models to the reduced
datasets. By associating an appropriate SkyModel, consisting of both temporal
and spectral components, or using custom temporal models with spectroscopic
variability, to each dataset, a joint fitting across the datasets will directly
return the best fit temporal and spectral parameters.

>>>>

A common use case in most astrophysical scenarios is to study the temporal
variability of the source. The most basic way to do this is to construct a
lightcurve, i.e., the flux of the source in each given time bin. In \gammapy, this
is done by using the \code{LightCurveEstimator} which fits the normalisation of the
source in each energy band per observation, keeping other parameters constant.
For custom time binning, an observation needs to be split into finer time bins using
the \code{Observation.select\_time} method. Figure~\ref{fig:hess_lightcurve_pks}
shows the lightcurve of the blazar PKS~2155-304 in different energy bands as
observed by the H.E.S.S. telescope during an exceptional flare on the night of
July 29 - 30, 2006~\cite{2009A&A...502..749A}. The data is available publicly
as a part of the HESS-DL3-DR1~\cite{HESS-DL3-DR1}, and shipped with
\verb"GAMMAPY_DATA". Each observation is first split into 10 min smaller
observations, and spectra extracted for each of these within a 0.11 deg radius
around the source. A \code{PowerLawSpectralModel} is fit to all the datasets, leading
to a reconstructed index of $3.54 \pm 0.02$. With this assumed spectral model
the \code{LightCurveEstimator} runs directly for two energy bands, 0.5 - 1.5
TeV, and 1.5 - 20 TeV, respectively.

@@

The obtained flux points can be analytically modelled using the available, or
user-implemented temporal models. Alternatively, instead of  extracting a
lightcurve, it is also possible to directly fit temporal models to the reduced
datasets. By associating an appropriate \code{SkyModel}, consisting of both temporal
and spectral components, or using custom temporal models with spectroscopic
variability, to each dataset, a joint fitting across the datasets will directly
return the best fit temporal and spectral parameters.
